---
title: "Figure1_plots"
author: "Miranda Hunter"
date: "5/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plots from Figure 1 of Hunter, Moncada et al.

Libraries:
``` {r}
library(Seurat)
library(tidyverse)
library(scales)
library(pals)
library(cowplot)
library(fgsea)
library(msigdbr)
library(data.table)
library(mvhspatialplots)
```

Load integrated object:
``` {r}
load("/Users/hunterm/Dropbox/MH_ST/Miranda_R/visium_Seurat_objs_fish/ABC_integrated_allgenes_new.R")
```

Apply metadata from Python clustering:
``` {r}
metadata <- read.csv('/Users/hunterm/Dropbox/MH_ST/200115_Visium_library_prep/Reuben analysis/Python_analyses/ABC_metadata_updated.csv')
ABC <- AddMetaData(ABC, 
                   metadata = metadata %>% dplyr::select(barcodes, tissue.type) %>% deframe(),
                   col.name = "tissue.type_new")
```

Figure 1d plots:
``` {r}
nice_spatial_dim_plot(ABC, group.by = "tissue.type_new")
```

Add UMAP embedding from Python clustering:
``` {r}
embeddings_new <- metadata %>% dplyr::select(barcodes, UMAP_1, UMAP_2) %>% column_to_rownames(var = "barcodes") %>% as.matrix()
rownames(embeddings_new) <- rownames(embeddings_new) %>% gsub(pattern = "\\.", replacement = "-", x = .)
ABC[["umap_new"]] <- CreateDimReducObject(embeddings = embeddings_new, key = "UMAP_", assay = "integrated")
```

Figure 1e plot:
``` {r}
DimPlot(ABC, reduction = "umap_new", group.by = "tissue.type_new")
```

Figure 1f plots: tissue type-specific genes.
``` {r}
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/markdown_manuscript_final/C1_new.R')

plots <- SpatialPlot(C1_new, 
            features = c("BRAFhuman", "actc1b", "tnnt2a", "mbpa"),
            alpha = c(0.1,1),
            image.alpha = 0,
            pt.size.factor = 1.3,
            stroke = 0,
            combine = F) 
plots <- lapply(plots, function(x) {
  x + scale_fill_gradientn(colours = rev(brewer.rdbu(n = 100)))
})
plot_grid(plotlist = plots, nrow = 1)
```

Figure 1g and S3a: spatial GO term pattening of tumor region
Load results from GO term spatial patterning analysis (see spatial_pattening_of_GO_terms_analysis.R):
``` {r}
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/go_term_spatial/A1/GO_expression_A1_tumor.R')
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/go_term_spatial/A1/GO_pval_A1_tumor.R')
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/go_term_spatial/B1/GO_expression_B1_tumor.R')
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/go_term_spatial/B1/GO_pval_B1_tumor.R')
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/go_term_spatial/C1/GO_expression_C1_tumor.R')
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/go_term_spatial/C1/GO_pval_C1_tumor.R')

load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/visium_Seurat_objs_fish/A1_norm.R')
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/visium_Seurat_objs_fish/B1_norm.R')
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/visium_Seurat_objs_fish/C1_norm.R')
```

Function to add spatial GO term info to metadata:
``` {r}
plot_spatial_GO_term <- function(pathway_add, plot_tumor = T, plot_interface = F) {
  # this assumes that the Seurat objects for ABC are already loaded, and the expression and pval matrices
  
  if (length(pathway_add) == 1) {
    
    pathway.add <- pathway_add
    samples <- c("A1", "B1", "C1")
    plots <- NULL
    
    if (plot_tumor) {
      for (sample in samples) {
        if (sample == "A1") {
          metadata.add <- A1.t.expr[rownames(A1.t.expr) %in% pathway.add,] %>% t() %>% rescale(to = c(-1,1)) %>% data.frame() %>% rownames_to_column() %>% deframe()
          names(metadata.add) <- gsub(pattern = "_1", replacement = "", x = names(metadata.add))
          A1 <- AddMetaData(A1, metadata = metadata.add, col.name = pathway.add)
          pval <- A1.t.pval[rownames(A1.t.pval) == pathway.add,] %>% signif(digits = 4)
          plots[[sample]] <- SpatialPlot(A1, features = pathway.add, image.alpha = 0, stroke = 0, pt.size.factor = 1.3) + 
            scale_fill_gradientn(colours = brewer.rdbu(n = 100) %>% rev()) +
            labs(title = pathway.add, subtitle = paste("P =", pval)) +
            theme(legend.position = "none", plot.title = element_text(size = 14, hjust = 0.5), plot.subtitle = element_text(size = 12, hjust = 0.5))
        } else if (sample == "B1") {
          metadata.add <- B1.t.expr[rownames(B1.t.expr) %in% pathway.add,] %>% t() %>% rescale(to = c(-1,1)) %>% data.frame() %>% rownames_to_column() %>% deframe()
          names(metadata.add) <- gsub(pattern = "_2", replacement = "", x = names(metadata.add))
          B1 <- AddMetaData(B1, metadata = metadata.add, col.name = pathway.add)
          pval <- B1.t.pval[rownames(B1.t.pval) == pathway.add,] %>% signif(digits = 4)
          plots[[sample]] <- SpatialPlot(B1, features = pathway.add, image.alpha = 0, stroke = 0, pt.size.factor = 1.3) + 
            scale_fill_gradientn(colours = brewer.rdbu(n = 100) %>% rev()) +
            labs(title = pathway.add, subtitle = paste("P =", pval)) +
            theme(legend.position = "none", plot.title = element_text(size = 14, hjust = 0.5), plot.subtitle = element_text(size = 12, hjust = 0.5))
        } else if (sample == "C1") {
          metadata.add <- C1.t.expr[rownames(C1.t.expr) %in% pathway.add,] %>% t() %>% rescale(to = c(-1,1)) %>% data.frame() %>% rownames_to_column() %>% deframe()
          names(metadata.add) <- gsub(pattern = "_3", replacement = "", x = names(metadata.add))
          C1 <- AddMetaData(C1, metadata = metadata.add, col.name = pathway.add)
          pval <- C1.t.pval[rownames(C1.t.pval) == pathway.add,] %>% signif(digits = 4)
          plots[[sample]] <- SpatialPlot(C1, features = pathway.add, image.alpha = 0, stroke = 0, pt.size.factor = 1.3) + 
            scale_fill_gradientn(colours = brewer.rdbu(n = 100) %>% rev()) +
            labs(title = pathway.add, subtitle = paste("P =", pval)) +
            theme(legend.position = "none", plot.title = element_text(size = 14, hjust = 0.5), plot.subtitle = element_text(size = 12, hjust = 0.5))
        }
      }
    } else if (plot_interface) {
      for (sample in samples) {
        if (sample == "A1") {
          metadata.add <- A1.i.expr[rownames(A1.i.expr) %in% pathway.add,] %>% t() %>% rescale(to = c(-1,1)) %>% data.frame() %>% rownames_to_column() %>% deframe()
          names(metadata.add) <- gsub(pattern = "_1", replacement = "", x = names(metadata.add))
          A1 <- AddMetaData(A1, metadata = metadata.add, col.name = pathway.add)
          pval <- A1.i.pval[rownames(A1.i.pval) == pathway.add,] %>% signif(digits = 4)
          plots[[sample]] <- SpatialPlot(A1, features = pathway.add, image.alpha = 0, stroke = 0, pt.size.factor = 1.3) + 
            scale_fill_gradientn(colours = brewer.rdbu(n = 100) %>% rev()) +
            labs(title = pathway.add, subtitle = paste("P =", pval)) +
            theme(legend.position = "none", plot.title = element_text(size = 14, hjust = 0.5), plot.subtitle = element_text(size = 12, hjust = 0.5))
        } else if (sample == "B1") {
          metadata.add <- B1.i.expr[rownames(B1.i.expr) %in% pathway.add,] %>% t() %>% rescale(to = c(-1,1)) %>% data.frame() %>% rownames_to_column() %>% deframe()
          names(metadata.add) <- gsub(pattern = "_2", replacement = "", x = names(metadata.add))
          B1 <- AddMetaData(B1, metadata = metadata.add, col.name = pathway.add)
          pval <- B1.i.pval[rownames(B1.i.pval) == pathway.add,] %>% signif(digits = 4)
          plots[[sample]] <- SpatialPlot(B1, features = pathway.add, image.alpha = 0, stroke = 0, pt.size.factor = 1.3) + 
            scale_fill_gradientn(colours = brewer.rdbu(n = 100) %>% rev()) +
            labs(title = pathway.add, subtitle = paste("P =", pval)) +
            theme(legend.position = "none", plot.title = element_text(size = 14, hjust = 0.5), plot.subtitle = element_text(size = 12, hjust = 0.5))
        } else if (sample == "C1") {
          metadata.add <- C1.i.expr[rownames(C1.i.expr) %in% pathway.add,] %>% t() %>% rescale(to = c(-1,1)) %>% data.frame() %>% rownames_to_column() %>% deframe()
          names(metadata.add) <- gsub(pattern = "_3", replacement = "", x = names(metadata.add))
          C1 <- AddMetaData(C1, metadata = metadata.add, col.name = pathway.add)
          pval <- C1.i.pval[rownames(C1.i.pval) == pathway.add,] %>% signif(digits = 4)
          plots[[sample]] <- SpatialPlot(C1, features = pathway.add, image.alpha = 0, stroke = 0, pt.size.factor = 1.3) + 
            scale_fill_gradientn(colours = brewer.rdbu(n = 100) %>% rev()) +
            labs(title = pathway.add, subtitle = paste("P =", pval)) +
            theme(legend.position = "none", plot.title = element_text(size = 14, hjust = 0.5), plot.subtitle = element_text(size = 12, hjust = 0.5))
        }
      }
    }
    spatial_plots <- plot_grid(plotlist = plots, nrow = 1)
    return(spatial_plots)
  }
}
```

``` {r}
p1 <- plot_spatial_GO_term("GO_MACROPHAGE_MIGRATION", plot_tumor = T, plot_interface = F)
p2 <- plot_spatial_GO_term("GO_SMALL_MOLECULE_BIOSYNTHETIC_PROCESS", plot_tumor = T, plot_interface = F)
p3 <- plot_spatial_GO_term("GO_EXTRACELLULAR_STRUCTURE_ORGANIZATION", plot_tumor = T, plot_interface = F)
plot_grid(p1,p2,p3, nrow = 3)
```

Figure 1h and S3b: spatial GO term pattening of microenvironment
``` {r}
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/go_term_spatial/A1_macroenv/GO_expression_A1_macroenv.R')
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/go_term_spatial/A1_macroenv/GO_pval_A1_macroenv.R')
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/go_term_spatial/B1_macroenv/GO_expression_B1_macroenv.R')
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/go_term_spatial/B1_macroenv/GO_pval_B1_macroenv.R')
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/go_term_spatial/C1_macroenv/GO_expression_C1_macroenv.R')
load('/Users/hunterm/Dropbox/MH_ST/Miranda_R/go_term_spatial/C1_macroenv/GO_pval_C1_macroenv.R')


p1 <- plot_spatial_GO_term("GO_MACROPHAGE_APOPTOTIC_PROCESS", plot_tumor = F, plot_interface = T)
p2 <- plot_spatial_GO_term("GO_SMALL_MOLECULE_BIOSYNTHETIC_PROCESS", plot_tumor = F, plot_interface = T)
p3 <- plot_spatial_GO_term("GO_EXTRACELLULAR_STRUCTURE_ORGANIZATION", plot_tumor = F, plot_interface = T)
plot_grid(p1,p2,p3, nrow = 3)
```



``` {r}
sessionInfo()
```

