---
title: "smalley_analyses"
author: "Miranda Hunter"
date: "7/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries:
``` {r, message = F, warning = F}
library(Seurat)
library(tidyverse)
library(scales)
library(pals)
library(cowplot)
library(fgsea)
library(msigdbr)
library(data.table)
library(mvhfunctions)
```


Read in counts matrix from GEO: https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE174401
``` {r}
smalley_counts <- Read10X_h5('/Users/hunterm/Dropbox/10x/Smalley_data/GSE174401_filtered_feature_bc_matrix.h5')
```

Make Seurat object:
``` {r}
smalley <- CreateSeuratObject(counts = smalley_counts,
                              project = "smalley",
                              min.cells = 3,
                              min.features = 200)
smalley
setwd('/Users/hunterm/Dropbox/10x/Smalley_data/')
save(smalley, file = "smalley_Seurat.R")
rm(smalley_counts)
```


QC metrics:
``` {r}
# add % mt genes to metadata
smalley[["percent.mt"]] <- PercentageFeatureSet(smalley, pattern = "^MT")

# plot some QC metrics
VlnPlot(smalley, features = c("nCount_RNA", "nFeature_RNA", "percent.mt"))
```

Filter cells with more than 20% MT reads as they did in the paper.
``` {r}
smalley <- subset(smalley, subset = percent.mt < 20)
smalley
```

Various processing steps:
``` {r, message = F, warning = F}
smalley <- SCTransform(smalley, verbose = F, vars.to.regress = "percent.mt")
smalley <- FindVariableFeatures(object = smalley, selection.method = "vst", verbose = F)
smalley <- ScaleData(object = smalley, features = rownames(smalley), verbose = F)
smalley <- RunPCA(object = smalley, features = VariableFeatures(smalley), verbose = F)
print(smalley[["pca"]], dims = 1:5, nfeatures = 10)
```

Select PCs for UMAP/clustering:
``` {r}
ElbowPlot(smalley)
```



## UMAP and clustering
``` {r}
smalley <- RunUMAP(smalley, dims = 1:15, verbose = F)
smalley <- FindNeighbors(smalley, dims = 1:15, verbose = F)
smalley <- FindClusters(smalley, verbose = T)
nice_dim_plot(smalley, group_by = "SCT_snn_res.0.8")

```

Look at some tumor genes:
``` {r}
tumor_genes <- c("MLANA", "PMEL", "DCT")
nice_feature_plot(smalley, features = tumor_genes, order = T, pt.size = 0.5)
```



Try to find "interface" cluster.
``` {r}
#load("/Users/hunterm/Dropbox/10x/R_objects/EF_integrated_allgenes_new.R")

Idents(EF.filt) <- "integrated_cell_type"
SC_interface_genes <- FindMarkers(EF.filt,
                                  ident.1 = "interface",
                                  assay = "SCT") %>%
  rownames_to_column(var = "gene") %>%
  filter(p_val_adj <= 0.05) %>%
  arrange(-avg_log2FC)

smalley <- AddModuleScore(smalley,
                              features = list(SC_interface_genes %>% filter(avg_log2FC >= 1.5) %>% pull(gene) %>% convert_to_human_list()),
                              name = "SC_interface")

smalley <- CellCycleScoring(smalley,
                                s.features = cc.genes$s.genes,
                                g2m.features = cc.genes$g2m.genes,
                                set.ident = T)

p1 <- nice_feature_plot(smalley, features = "SC_interface1", diverging_cmap = T, order = T)
p2 <- nice_dim_plot(smalley, group_by = "Phase", label = F)
plot_grid(p1,p2)
```



Score cells for tumor genes, ETS genes and cilia genes.
``` {r}
tumor_genes <- c("MLANA", "PMEL", "DCT", "MITF")
cilia_genes <- readxl::read_excel("/Users/hunterm/Dropbox/MH_ST/Miranda_R/cilia/cilia_gene_list_full.xlsx") %>% pull("Gene Name") %>% unique()
ets_genes <- c("ETS1", "ETS2", "ETV6", "ETV7", "ELF3", "EHF", "ELF5", "SPDEF", "GABPA", "KCNH6", "FEV", "ELK1", "EPHB1", "ELF1", "ELF2", "ELF4", "ETV4", "ETV5", "ETV1", "ERF", "ETV3", "SPI1", "SPIB", "SPIC", "ETV2") 

gene_lists <- list(tumor = tumor_genes,
                   cilia = cilia_genes,
                   ets = ets_genes)


for (ii in 1:length(gene_lists)) {
  marker_genes <- intersect(gene_lists[[ii]], rownames(smalley))
  marker_name <- names(gene_lists[ii])
  smalley <- AddModuleScore(smalley,
                            features = list(marker_genes),
                            name = marker_name)
}

nice_feature_plot(smalley, features = "tumor1", order = T)
nice_feature_plot(smalley, features = "cilia1", order = T)
nice_feature_plot(smalley, features = "ets1", order = T)
nice_feature_plot(smalley, features = "SC_interface1", order = T)

```

Add patient information (seems to be stored in the cell barcode):
``` {r}
patient_info <- smalley[[]] %>% rownames_to_column(var = "cell") %>% dplyr::select(cell)

patient_info$pt <- gsub(pattern = ".*-",
                        x = patient_info$cell,
                        replacement = "")

smalley <- AddMetaData(smalley,
                       metadata = deframe(patient_info),
                       col.name = "patient")
nice_dim_plot(smalley, group_by = "patient", label = F)
```

Add some cluster info.
``` {r}
smalley <- FindClusters(smalley, resolution = 0.4, verbose = F)
smalley <- FindClusters(smalley, resolution = 0.3, verbose = F)
nice_dim_plot(smalley, group_by = "SCT_snn_res.0.8")
nice_dim_plot(smalley, group_by = "SCT_snn_res.0.4")
nice_dim_plot(smalley, group_by = "SCT_snn_res.0.3")
nice_feature_plot(smalley, features = "tumor1", order = T)
nice_feature_plot(smalley, features = "cilia1", order = T)
nice_feature_plot(smalley, features = "ets1", order = T)
nice_feature_plot(smalley, features = "SC_interface1", order = T)
```


Add cluster annotations:
``` {r}
smalley_metadata <- smalley[[]] %>% rownames_to_column(var = "cell") %>% dplyr::select(cell, SCT_snn_res.0.3)
for (ii in 1:nrow(smalley_metadata)) {
  cluster <- smalley_metadata[ii,2] %>% as.character() %>% as.numeric()
  if (cluster == 11) {
    smalley_metadata$cell_type[ii] <- "pDC"
  } else if (cluster %in% c(6,14,8,2,9,5,7)) {
    smalley_metadata$cell_type[ii] <- "tumor"
  } else if (cluster %in% c(3,13)) {
    smalley_metadata$cell_type[ii] <- "monocytes, macrophages, DCs"
  } else if (cluster %in% c(0,1,4,16)) {
    smalley_metadata$cell_type[ii] <- "T/NK cells"
  } else if (cluster == 10) {
    smalley_metadata$cell_type[ii] <- "B cells"
  } else if (cluster == 15) {
    smalley_metadata$cell_type[ii] <- "fibroblasts"
  } else if (cluster == 12) {
    smalley_metadata$cell_type[ii] <- "neurons"
  } else if (cluster == 17) {
    smalley_metadata$cell_type[ii] <- "unknown"
  } 
}

smalley <- AddMetaData(smalley,
                   metadata = smalley_metadata %>% dplyr::select(cell, cell_type) %>% deframe(),
                   col.name = "cell_type")

nice_dim_plot(smalley, group_by = "cell_type")
```

Separate out "interface" cells.

Classify cells as "interface":
``` {r}
interface_cutoff <- 0.5

interface_cells <- smalley[[]] %>% rownames_to_column(var = "cell") %>% filter(SC_interface1 > interface_cutoff) %>% dplyr::select(cell) %>% mutate(cell_type = "interface") %>% deframe()
other_cells <- smalley[[]] %>% rownames_to_column(var = "cell") %>% .[!.$cell %in% names(interface_cells),] %>% dplyr::select(cell, cell_type) %>% deframe()

# add new metadata column labelling interface
smalley <- AddMetaData(smalley,
                           metadata = c(interface_cells, other_cells),
                           col.name = "interface_cell")

smalley <- AddMetaData(smalley,
                       metadata = interface_cells,
                       col.name = "interface")

nice_dim_plot(smalley, group_by = "interface_cell", label = F)
nice_dim_plot(smalley, group_by = "interface", label = F)

Idents(smalley) <- "interface_cell"
nice_dim_plot(subset(smalley, idents = "interface"), group_by = "cell_type")
```

Break the interface into different cell types:
``` {r}
smalley_metadata <- smalley[[]] %>% rownames_to_column(var = "cell") %>% dplyr::select(cell, cell_type, interface_cell)
interface_metadata <- smalley_metadata %>% filter(interface_cell == "interface") # 515 interface cells

interface_metadata$interface_cell_type <- paste0("interface (", interface_metadata$cell_type, "-like)")

new_metadata <- c(interface_metadata %>% dplyr::select(cell, interface_cell_type) %>% deframe(),
                  smalley_metadata %>% filter(interface_cell != "interface") %>% dplyr::select(cell, cell_type) %>% deframe())

smalley <- AddMetaData(smalley,
                       metadata = new_metadata,
                       col.name = "interface_cell_type")
nice_dim_plot(smalley, group_by = "interface_cell_type")
```


Make plots for figure:
``` {r}
p1 <- DimPlot(smalley, group.by = "cell_type", cols = tol(8), pt.size = 0.75)
p2 <- FeaturePlot(smalley, features = "SC_interface1", max.cutoff = 1, order = T, pt.size = 0.75) + scale_color_gradientn(colours = rev(magma(n = 100)))
p3 <- DimPlot(smalley, group.by = "interface", pt.size = 0.75)
plot_grid(p1,p2,p3, align = "hv", axis = "lrbt", ncol = 3)

```

Feature plots:
``` {r}
genes <- c("PLK1", "HMGB2", "TUBB4B", "TPX2")

plots <- FeaturePlot(smalley,
            features = genes,
            #min.cutoff = -2,
            #max.cutoff = 2,
            pt.size = 0.2,
            #slot = "scale.data",
            order = F,
            combine = F)
plots <- lapply(plots, function(x) {
  x + scale_color_gradientn(colors = viridis(n = 100))
})

plot_grid(plotlist = plots, nrow = 2, align = "hv", axis = "lrbt")

plots <- FeaturePlot(smalley,
            features = genes,
            #min.cutoff = -2,
            #max.cutoff = 2,
            pt.size = 0.4,
            #slot = "scale.data",
            order = T,
            combine = F)
plots <- lapply(plots, function(x) {
  x + scale_color_gradientn(colors = viridis(n = 100))
})

plot_grid(plotlist = plots, nrow = 2, align = "hv", axis = "lrbt")
```

Cilia expression:
``` {r}

# tumor
plots1 <- VlnPlot(object = subset(smalley, idents = c("tumor", "interface (tumor-like)")),
              features = "cilia1",
              group.by = "interface_cell_type",
              sort = F,
              combine = F)
plots1 <- lapply(plots1, function(x) {
  x + theme(axis.title.x = element_blank(), axis.text.x = element_blank()) + NoLegend() + geom_hline(yintercept = 0, linetype = "dashed")
})
p1 <- plot_grid(plotlist = plots1)

# T/NK
plots2 <- VlnPlot(object = subset(smalley, idents = c("T/NK cells", "interface (T/NK cells-like)")),
              features = "cilia1",
              group.by = "interface_cell_type",
              sort = F,
              combine = F)
plots2 <- lapply(plots2, function(x) {
  x + theme(axis.title.x = element_blank(), axis.text.x = element_blank()) + NoLegend() + geom_hline(yintercept = 0, linetype = "dashed")
})
p2 <- plot_grid(plotlist = plots2)

# myeloid
plots3<- VlnPlot(object = subset(smalley, idents = c("monocytes, macrophages, DCs", "interface (monocytes, macrophages, DCs-like)")),
              features = "cilia1",
              group.by = "interface_cell_type",
              sort = F,
              combine = F)
plots3 <- lapply(plots3, function(x) {
  x + theme(axis.title.x = element_blank(), axis.text.x = element_blank()) + NoLegend() + geom_hline(yintercept = 0, linetype = "dashed")
})
p3 <- plot_grid(plotlist = plots3)

plot_grid(p1,p2,p3, nrow = 3, align = "hv", axis = "lrbt")
```



Stats:
tumor:
``` {r}
tumor_metadata <- smalley[[]] %>% filter(cell_type == "tumor") %>% dplyr::select("cilia1", "ets1", "interface_cell_type") 

stats <- wilcox.test(tumor_metadata %>% filter(interface_cell_type == "tumor") %>% pull(cilia1),
            tumor_metadata %>% filter(interface_cell_type == "interface (tumor-like)") %>% pull(cilia1))
stats$p.value
```


T/NK:
``` {r}
metadata <- smalley[[]] %>% filter(cell_type == "T/NK cells") %>% dplyr::select("cilia1", "ets1", "interface_cell_type") 

stats <- wilcox.test(metadata %>% filter(interface_cell_type == "T/NK cells") %>% pull(cilia1),
            metadata %>% filter(interface_cell_type == "interface (T/NK cells-like)") %>% pull(cilia1))
stats$p.value
```


Myeloid:
``` {r}
metadata <- smalley[[]] %>% filter(cell_type == "monocytes, macrophages, DCs") %>% dplyr::select("cilia1", "ets1", "interface_cell_type") 

stats <- wilcox.test(metadata %>% filter(interface_cell_type == "monocytes, macrophages, DCs") %>% pull(cilia1),
            metadata %>% filter(interface_cell_type == "interface (monocytes, macrophages, DCs-like)") %>% pull(cilia1))
stats$p.value
```


Violin plot of interface scores:
``` {r}
#smalley_for_plotting <- subset(smalley, idents = c("interface (T/NK cells-like)", "interface (monocytes, macrophages, DCs-like)", "interface (tumor-like)", "neurons", "tumor", "pDC", "T/NK cells", "B cells", "unknown", "monocytes, macrophages, DCs", "fibroblasts"))

Idents(smalley_for_plotting) <- "interface_cell_type"
levels(smalley_for_plotting) <- c("interface (tumor-like)", "interface (T/NK cells-like)", "interface (monocytes, macrophages, DCs-like)", "tumor", "T/NK cells", "monocytes, macrophages, DCs", "neurons", "B cells", "pDC", "fibroblasts", "unknown")
nice_violin_plot(smalley_for_plotting, features = "SC_interface1", pt.size = 0, sort = F)
```

stats:
``` {r}
smalley_metadata <- smalley[[]]

stats <- pairwise.wilcox.test(smalley_metadata$SC_interface1, smalley_metadata$interface_cell_type, p.adjust.method = "bonferroni")
stats$p.value
```


