---
title: "Fig2 calculation of interface enriched genes"
author: "Miranda Hunter"
date: "6/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries:
``` {r, message = F, warning = F}
library(Seurat)
library(tidyverse)
library(scales)
library(pals)
library(cowplot)
library(fgsea)
library(msigdbr)
library(data.table)
library(mvhfunctions)
```


Load SRT dataset.
``` {r}
load("/Users/hunterm/Dropbox/MH_ST/Miranda_R/visium_Seurat_objs_fish/ABC_integrated_allgenes_new.R")
Idents(ABC) <- "cell_type"
```


1. Calculate avg log2FC of each gene in the dataset within the interface region, relative to the tumor and muscle clusters.
``` {r}
interface_markers <- FindMarkers(ABC,
                                 ident.1 = "interface",
                                 #ident.2 = c("tumor", "muscle"),
                                 logfc.threshold = 0,
                                 assay = "SCT") %>%
  rownames_to_column(var = "gene") %>%
  arrange(-avg_log2FC) 

```


2. Calculate avg log2FC of each of the genes above within the tumor and muscle clusters.
Tumor:
``` {r}
interface_markers_tumor <- FindMarkers(ABC,
                                       ident.1 = "tumor",
                                       features = interface_markers$gene,
                                       assay = "SCT",
                                       logfc.threshold = 0) %>% 
  rownames_to_column(var = "gene") %>%
  arrange(-avg_log2FC)
```

Muscle:
``` {r}
interface_markers_muscle <- FindMarkers(ABC,
                                        ident.1 = "muscle",
                                        features = interface_markers$gene,
                                        assay = "SCT",
                                        logfc.threshold = 0) %>% 
  rownames_to_column(var = "gene") %>%
  arrange(-avg_log2FC)
```


3. Determine which genes have a higher log2FC within the interface than in the tumor or muscle.
``` {r}
interface_markers <- interface_markers %>% dplyr::select(gene, avg_log2FC, p_val_adj)
interface_markers_tumor <- interface_markers_tumor %>% dplyr::select(gene, avg_log2FC)
interface_markers_muscle <- interface_markers_muscle %>% dplyr::select(gene, avg_log2FC)

colnames(interface_markers)[2] <- "log2FC_in_interface"
colnames(interface_markers_tumor)[2] <- "log2FC_in_tumor"
colnames(interface_markers_muscle)[2] <- "log2FC_in_muscle"

all_expression_data <- merge(x = interface_markers,
                             y = interface_markers_tumor,
                             by = "gene") %>%
  merge(x = .,
        y = interface_markers_muscle,
        by = "gene")


all_expression_data_filt <- all_expression_data %>% filter(p_val_adj <= 0.05)
```

Determine upregulated genes as genes with an interface log2FC > 0, and genes with a higher interface log2FC than tumor AND muscle log2FCs.
``` {r}
interface_genes_up <- all_expression_data_filt %>% filter(log2FC_in_interface > 0 & log2FC_in_interface > log2FC_in_muscle & log2FC_in_interface > log2FC_in_tumor)
interface_genes_down <- all_expression_data_filt %>% filter(log2FC_in_interface < 0 & log2FC_in_interface < log2FC_in_muscle & log2FC_in_interface < log2FC_in_tumor)
```


4. Plot volcano plot of log2FC and adjusted p vals in interface, and label the interface marker genes calculated in step 3.
``` {r}

interface_genes_up_or_down <- c(interface_genes_up$gene, interface_genes_down$gene)

all_expression_data_plot <- rbind(all_expression_data[all_expression_data$gene %in% interface_genes_up$gene,] %>% mutate(interface = "up"),
                                  all_expression_data[all_expression_data$gene %in% interface_genes_down$gene,] %>% mutate(interface = "down"),
                                  all_expression_data[!all_expression_data$gene %in% interface_genes_up_or_down,] %>% mutate(interface = NA))

ggplot(all_expression_data_plot, aes(x = log2FC_in_interface, y = -log10(p_val_adj), color = interface)) +
  geom_point() +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) + 
  #coord_cartesian(xlim = c(-2,1)) +
  theme_minimal() +
  xlab("average log2 fold change in interface") +
  ylab("-log10 p-value") +
  theme(axis.title = element_text(size = 16, color = "black"),
        axis.text = element_text(size = 16, color = "black")) +
  geom_text_repel(data = subset(all_expression_data_plot, gene %in% interface_genes_up_or_down), aes(label = gene), colour = "black",
                  min.segment.length = 0.05)

```














